---
title: "Code Review - stat_summary"
author: "Nick Lesniak"
date: "`r Sys.Date()`"
output: 
  github_document: default
  html_document: default
---


Using data from COVID Racial Data Tracker https://covidtracking.com/  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = T)

covid_data <- read_csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vS8SzaERcKJOD_EzrtCDK1dX1zkoMochlA9iHoHg_RSw3V8bkpfk1mpw4pfL5RdtSOyx_oScsUtyXyk/pub?gid=43720681&single=true&output=csv',
         col_types = cols(.default = col_double(), State = col_character())) %>% 
  pivot_longer(cols = -c(State, Date), names_to = 'data_type', values_to = 'value') %>% 
  filter(!grepl('Ethni', data_type)) %>% 
  mutate(data_type = gsub('Ethnicity_', '', data_type),
         Date = lubridate::as_date(
            gsub('(\\d{4})(\\d{2})(\\d{2})', '\\1-\\2-\\3', Date))) %>% 
  separate(data_type, c('type', 'race_ethnicity'))

covid_totals <- covid_data %>% 
  filter(race_ethnicity == 'Total') %>% 
  rename(total = value) %>% 
  select(-race_ethnicity)

covid_data <- covid_data %>% 
  filter(race_ethnicity != 'Total',
         type %in% c('Cases', 'Deaths')) %>% 
  left_join(covid_totals, by = c('Date', 'State', 'type')) %>% 
  mutate(per100people = round(value/total * 100, 0))

mi_cases <- covid_data %>% 
  filter(State == 'MI', 
         type == 'Cases',
         !is.na(per100people))
  
```

### Michigan COVID cases by race from April 12, 2020 through March 7, 2021

```{r points}
mi_cases %>% 
  ggplot(aes(x = race_ethnicity, y = per100people)) + 
    geom_point(alpha = 0.1, position = position_jitter(0.3)) + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```

This causes a lot of overplotting, let's calculate the summary statistics and plot the median and interquartile range.


```{r calc_median_iqr}
mi_cases %>% 
  group_by(type, race_ethnicity) %>% 
  mutate(median = median(per100people),
         lower = quantile(per100people, probs = c(0.25), na.rm = T),
         upper = quantile(per100people, probs = c(0.75), na.rm = T)) %>% 
  ggplot(aes(x = race_ethnicity)) + 
    geom_point(aes(y = median)) + 
    geom_errorbar(aes(ymin = lower, ymax = upper)) + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```

Creating the plot with mutate works, but...
what if we want to change anything in the plot?  
  
  
What would need to be changed?    
  
  

What if we could set the plotted stats in the plot itself?

### stat_summary

You may have actually already been using stat_summary - for example, `geom_bar(fun = "mean", stat = "summary")` is using `stat_summary`.  
`stat_summary` allows us to visualize any summary statistic with most geoms. It is similar to grouping and summarizing within the ggplot layer

```{r}
mi_cases %>% 
  group_by(race_ethnicity) %>% 
  summarize(y = median(per100people))
```


Let's use the `stat_summary` function to tidy up our plotting code

```{r point plot}
mi_cases %>% 
  ggplot(aes(x = race_ethnicity, y = per100people)) + 
    stat_summary(fun = median, 
                 geom = 'point') + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```

Here we specify the function to be applied to y, `fun = median`, and how we want it visualized, `geom = 'point'`. With this information, the median will be plotted based on the plotting groups, no need to recalcuate if changing groupings. But what if we wanted to add in error bars?


We can specify the min and max for geoms that can plot that information. We can use R functions.


```{r range plot}
mi_cases %>% 
  ggplot(aes(x = race_ethnicity, y = per100people)) + 
    stat_summary(fun = median,
                 fun.min = min,
                 fun.max = max,
                 geom = 'pointrange') + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```

Or we can input our own functions.


```{r plot}
mi_cases %>% 
  ggplot(aes(x = race_ethnicity, y = per100people)) + 
    stat_summary(fun = median,
                 fun.min = function(x)quantile(x, probs = 0.25, na.rm = T),
                 fun.max = function(x)quantile(x, probs = 0.75, na.rm = T),
                 geom = 'pointrange') + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```


But for something like mean and confidence intervals, no need to create this ourselves, theres a function for that.
We can use functions from the `Hmisc` package to process our data using `fun.data`, this takes the data and calculates the `y`, `ymin` and `ymax`, then outputs a dataframe from ggplot to plot.


```{r}
median_hilow
```


To see how these functions work, we need to look at the Hmisc functions that hmisc/ggplot are referencing  
`?Hmisc::smedian.hilow`

```{r}
Hmisc::smedian.hilow
```


```{r}
mi_cases %>% 
  group_by(race_ethnicity) %>% 
  summarise(median_hilow(per100people))
```

Others functions are  
`mean_cl_boot()`  
`mean_cl_normal()`  
`mean_sdl()`  
`median_hilow()` 


```{r}
mi_cases %>% 
  ggplot(aes(x = race_ethnicity, y = per100people)) + 
    	stat_summary(fun.data = median_hilow, 
    	             fun.args = list(conf.int=0.5)) +
      theme_bw() + labs(x = NULL, y = 'Cases per 100 people')

```

In this example we specified `fun.args = list(conf.int=0.5)` to give us the interquartile range. 


Using this function we can change the data we are plotting without the need to recalculate and make another new dataframe.
Let's plot the cases of COVID over time by race for all states.

### Cases of COVID over time by race for all states.

```{r}
covid_data %>% 
  filter(type == 'Cases') %>% 
  ggplot(aes(x = Date, y = per100people, color = race_ethnicity)) + 
    stat_summary(fun.data = median_hilow, 
  	             fun.args = list(conf.int=0.5)) + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people', color = NULL)

```

We can customize our plot by layering different `stat_summary` layers.


```{r}
covid_data %>% 
  filter(type == 'Cases',
         !is.na(per100people)) %>% 
  ggplot(aes(x = Date, y = per100people, color = race_ethnicity, fill = race_ethnicity)) + 
    stat_summary(fun = median, geom = 'line') + 		
    stat_summary(fun.data = median_hilow, fun.args = list(conf.int = 0.5), 
                 geom = 'ribbon', alpha = 0.1, size = 0.1) + 
    theme_bw() + labs(x = NULL, y = 'Cases per 100 people', color = NULL)

```

### Your turn, what do you see in the COVID data for your home state/or other region of interest?

```{r, eval = FALSE}
covid_data %>% 

```

#### OptiFit project

Kelly & Nick's breakout room

```{r}
library(glue)
optifit <- read_tsv('optifit_results.tsv')
optifit %>% 
  mutate(strategy = case_when(
    strategy == "database" ~ as.character(glue("database_{ref}")),
    TRUE ~ strategy)) %>% 
  mutate(strategy = fct_reorder(as.factor(strategy), mcc)) %>% 
  ggplot(aes(x = strategy, y = mcc, color = dataset)) + 
    stat_summary(fun.data = median_hilow, 
  	             fun.args = list(conf.int=0.5)) +
  facet_wrap('dataset') +
  coord_flip() +
  theme_bw()
```

